---
title: "GLMM_Handbook"
author: "Elizaveta Shcherbinina"
date: "2025-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(MuMIn)
library(broom.mixed)
library(ggplot2)
library(dplyr)
library(tidyr)
library(sjPlot)
```

# Data Exploration 

```{r glimpse on data}
data <- read.csv("data/202509011_landindex_clim_poll_soil_fsize.csv")
data <- data[,-1]
glimpse(data)
```


```{r lrr calculation}
# Fill only the missing values in your existing columns
#df$LRR <- ifelse(is.na(df$LRR), df_es$yi, df$LRR)
#df$LRR_vi <- ifelse(is.na(df$LRR_vi), df_es$vi, df$LRR_vi)
```


```{r years vs yield}
plot(data$mean_yield_control_kgha~data$landcover_map_year)
plot(data$mean_yield_treatment_kgha~data$landcover_map_year)

plot(data$LRR~data$landcover_map_year)
```

```{r was anything else changing}

plot(data$nat.hab.1000~data$landcover_map_year) # no
plot(data$cropland.1000~data$landcover_map_year) # no
plot(data$ocs_0.30cm_mean.1000~data$landcover_map_year) # whats that ?? 

```

```{r precipitation and temperature}
plot(data$prec_avg_1970.2000~data$landcover_map_year) # more or less the same
plot(data$temp_avg_1970.2000~data$landcover_map_year) # more or less the same 

```

```{r crop types and treatments}
data %>% 
  ggplot(aes(x = crop_type_grouped_small, fill = treatment)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Crop type", y = "Share of treatments", fill = "Treatment") +
  theme_minimal()
```

```{r croptypes and pollination}
data %>% 
  select(crop_type, crop_type_grouped_small, crop_type_grouped_big, poll_dependent) %>% 
  unique() # grapes are separated for non-pollinated fruits 
```

```{r nat hab summary}
summary(data$nat.hab.1000)
```

```{r regularity of LRR}
data %>%
 # filter(country_new == "China") %>% 
 # mutate(habitat = ifelse(nat.hab.1000 > 0.5, "natural", "industrial")) %>%
  group_by(landcover_map_year, fieldsize) %>%
  summarise(mean_LRR = mean(LRR, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = landcover_map_year, y = mean_LRR, color = fieldsize)) +
  geom_line() +
  geom_point() +~
  labs(x = "Year", y = "Mean LRR", color = "Habitat") +
  theme_minimal() 

# I m really interested what happened in 2003 (?)
# ALso, the verysmall area is indeed the most stable one
```



# GLMM 

```{r sclaing}

scale_df <- function(df) {
  numeric_cols <- sapply(df, is.numeric)
  df[numeric_cols] <- lapply(df[numeric_cols], scale)
  return(df)}

df <- scale_df(data)

```


Effects to check: 
- treatment 
- precipitation 
- temperature
- croptype
- pollination 
- study_pubyear

landscape: 
- complexity
- composition
- fieldsize 

ecosystem sevices 
- nat.hab -> provisioning 
- SR species richness 

soil:

environmental:
- precipitation 
- temperature
- latitude 

```{r correlation}
# didn't finish, please proceed
numeric_cols <- sapply(df, is.numeric)
df_numeric <- df[numeric_cols]
df_numeric <- 
df_numeric %>% 
  select(-harvest_year, -publication_to_harvest_year_difference,
         -yield_SD_control, -yield_SD_treatment,
         -LRR_vi, -small_area_m.5000,
         -small_area_m.2500, -small_area_m.1000,
         -large_area_m.5000, -large_area_m.2500,
         -large_area_m.1000, -verylarge_area_m.2500,
         -verylarge_area_m.1000, -medium_area_m.5000,
         -medium_area_m.2500, -medium_area_m.1000,
         -medium_area_m.2500, -medium_area_m.1000)
  
df_env <- 
df %>% 
  select(LRR, prec_avg_1970.2000, temp_avg_1970.2000, latitude_decimal)

cor_mat <- rstatix::cor_mat(df_env)
```


```{r land uses effect}

nathab1000 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + nat.hab.1000, data = df)  
nathabwograss1000 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + nat.hab.wo.grass.1000, data = df)  
cropland1000 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + cropland.1000, data = df)  
```

```{r}
library(splines)

model <- lmer(LRR ~ (1|ma_id) + (1|study_id) + prec_avg_1970.2000 + ns(latitude_decimal, df = 4), data = df)

preclat <- lmer(LRR ~ (1|ma_id) + (1|study_id) + prec_avg_1970.2000 + latitude_decimal, data = df)  

plot_model(preclat, type = "slope") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),        # removes default title
    plot.subtitle = element_text(size = 12)
  ) 

preclat <- lmer(LRR ~ (1|ma_id) + (1|study_id) + + prec_avg_1970.2000+ latitude_decimal, data = df)  
```


```{r}
# Generate predictions
library(ggeffects)
library(ggplot2)
pred <- ggpredict(model, terms = "prec_avg_1970.2000")
plot(pred) + theme_minimal()

```



```{r}
glimpse(df)
```

```{r}
# Model 2: More complex model
model2 <- lmer(LRR ~ nat.hab.1000 + clay_0.5cm_mean.1000 + soc_0.5cm_mean.1000 + crop_type_grouped_big +
                 (1 + nat.hab.1000 | ma_id), # Random intercept and slope for ma_id
               data = df,
               REML = FALSE) # Use ML if you plan to compare fixed effects using likelihood ratio tests

summary(model2)
```

```{r}
library(DHARMa)
library(performance) 
check_collinearity(model2)

# Simulate residuals with DHARMa
sim_resids <- simulateResiduals(fittedModel = model2)
plot(sim_resids)
```

```{r correlation plot}

library(dplyr)
library(corrplot)

df_numeric <- df %>%
  select(
    temp_avg_1970.2000,
    prec_avg_1970.2000,
    nofield_area_m.1000,
    small_area_m.1000,
    verysmall_area_m.1000,
    large_area_m.1000,
    verylarge_area_m.1000,
    medium_area_m.1000,
    SR,
    soc_15.30cm_mean.1000,
    ocs_0.30cm_mean.1000,
    clay_15.30cm_mean.1000,
    simpsonsevenness.1000,
    shannon.1000,
    inert.edgelength.1000,
    inert.peri.area.ratio.1000,
    inert.1000,
    crop.edgelength.1000,
    crop.peri.area.ratio.1000,
    cropland.1000,
    nat.hab.wo.grass.edgelength.1000,
    nat.hab.wo.grass.peri.area.ratio.1000,
    nat.hab.wo.grass.1000,
    nat.hab.peri.area.ratio.1000,
    nat.hab.edgelength.1000,
    nat.hab.1000,
    latitude_decimal,
    landcover_map_year
  ) %>%
  mutate(across(where(is.matrix), as.vector)) %>%
  mutate(across(everything(), as.numeric))

cor_matrix <- cor(df_numeric, use = "pairwise.complete.obs")

print(cor_matrix)


corrplot(cor_matrix,
         method = "color",
         type = "upper",
         tl.cex = 0.8,
         title = "Pairwise Complete Correlation")
```


```{r}
model_1 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + SR + nat.hab.1000, data = df, REML = TRUE)
summary(model_1)

model_2 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + nat.hab.peri.area.ratio.1000 + crop.peri.area.ratio.1000 , data = df, REML = TRUE)
summary(model_2)

model_21 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + nat.hab.peri.area.ratio.1000 + crop.peri.area.ratio.1000 +SR, data = df, REML = TRUE)
summary(model_21)


model_3 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + soc_30.60cm_mean.1000 + clay_15.30cm_mean.2500 + latitude_decimal, data = df, REML = TRUE)
summary(model_3)

model_4 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + SR + shannon.1000  + soc_30.60cm_mean.1000 + clay_15.30cm_mean.2500 + latitude_decimal, data = df, REML = TRUE)
summary(model_4)

model_4 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + SR + shannon.1000  + ocs_0.30cm_mean.1000 + clay_15.30cm_mean.2500, data = df, REML = TRUE)
summary(model_4)

plot_model(model_4, type = "slope") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),        # removes default title
    plot.subtitle = element_text(size = 12)
  ) +
  labs(subtitle = "Intercept is fieldsize = large 
AIC = 552.3438")


```

```{r}
# For model_1: LRR ~ SR + nat.hab.1000
plot(ggpredict(model_1, terms = "SR")) + labs(title = "Effect of Species Richness (SR) on LRR")
plot(ggpredict(model_1, terms = "nat.hab.1000")) + labs(title = "Effect of Natural Habitat (1000m) on LRR")
```
```{r}
# For model_1: LRR ~ SR + nat.hab.1000
plot_model(model_1, type = "pred", terms = "SR")
plot_model(model_1, type = "pred", terms = "nat.hab.1000")
```
## START HERE


### H1: SR + nat.hab.peri.area.ratio.1000 + poll_dependent
```{r}
#H1: species richness will be travelling through the edges of natural landcsapes 
# and hence increasing the pollinated and not affecting the not pollinated
# the effect on both can also be damaging because of the pest 

df$poll_dependent <- as.factor(df$poll_dependent)

hyp1 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + SR + nat.hab.peri.area.ratio.1000 + poll_dependent, data = df, REML = TRUE)

summary_hyp1 <- summary(hyp1)
AIC_hyp1 <- AIC(hyp1)

plot_model(hyp1, type = "slope") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(), 
    plot.subtitle = element_text(size = 12)) +
  labs(subtitle = "Intercept is fieldsize = large 
AIC = 3647.199")

```

```{r}
plot_model(hyp1, type = "pred", terms = "SR")
plot_model(hyp1, type = "int", terms = c("SR", "nat.hab.peri.area.ratio.1000"))
plot_model(hyp1, type = "re")     # random effects
# A288: treat organic vs contr conventional
# D352:	Erosion simulation

data %>% 
  filter(ma_id == "D352")
```
Here we see that the response to nat.hab.prei.area.ratio.1000 and SR nonlinear are 

There are several methods to transform nonlinear responses: 
- poly
```{r poly error}
#Error in poly(dots[[1L]], degree, raw = raw, simple = raw && nd > 1) : 
# missing values are not allowed in 'poly'

#hyp1_poly <- lmer(LRR ~ (1|ma_id) + (1|study_id) + poly(SR, 2) + poly(nat.hab.peri.area.ratio.1000, 2) + poll_dependent, data = df, REML = TRUE)

```

- equation
for nat.hab.peri.area I cannot imagine an equations that would suit the function so i ll proceed to spines
- splines

```{r}
library(splines)

hyp1_splines <- lmer(LRR ~ (1|ma_id) + (1|study_id) + ns(nat.hab.peri.area.ratio.1000, df = 3) + ns(SR, df = 2) + poll_dependent, data = df)

summary_hyp1_splines <- summary(hyp1_splines)
AIC_hyp1_splines <- AIC(hyp1_splines)

plot_model(hyp1_splines, type = "est") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),        
    plot.subtitle = element_text(size = 12)
  ) 

```


```{r}
library(splines)

hyp1_randomslope <- lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|latitude_decimal) + ns(nat.hab.peri.area.ratio.1000, df = 3) + ns(SR, df = 2) + poll_dependent, data = df)

summary(hyp1_randomslope)
AIC_hyp1_splines <- AIC(hyp1_splines)

plot_model(hyp1_randomslope, type = "slope") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 

```

(... + nat.hab.1000 | ma_id): Allow the strength and direction (slope) of the relationship between nat.hab.1000 and LRR to be different for each ma_id group. It accounts for the possibility that the effect of natural habitat on the yield response might be stronger in some regions (ma_id) and weaker in others.

```{r visualisation ideas}
library(effects)
library(ggplot2)
plot(predictorEffects(hyp1, ~ SR + nat.hab.peri.area.ratio.1000 + poll_dependent),
     multiline = TRUE, ci.style = "bands")
```

```{r}
hyp1_interaction <- lmer(LRR ~ (1|ma_id) + (1|study_id) + ns(nat.hab.peri.area.ratio.1000, df = 3) * ns(SR, df = 2), data = df)
plot_model(hyp1_interaction, type = "est") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 
```


### H2: (1|ma_id) + (1|study_id) + nat.hab.peri.area.ratio.1000 + crop.peri.area.ratio.1000


```{r}
hyp2 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + nat.hab.peri.area.ratio.1000+crop.peri.area.ratio.1000 , data = df, REML = TRUE)
summary(hyp2)
```




### H3: (1|ma_id) + (1|study_id) + soc_30.60cm_mean.1000 + clay_15.30cm_mean.2500 + latitude_decimal

```{r}
hyp3 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + soc_30.60cm_mean.1000 + clay_15.30cm_mean.2500 + latitude_decimal, data = df, REML = TRUE)
summary(model_3)
```

### H4: Ecological Effects
(1|ma_id) + (1|study_id) + SR + shannon.1000  + soc_30.60cm_mean.1000 + clay_15.30cm_mean.2500 + latitude_decimal

```{r}
hyp4 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + SR + shannon.1000  + soc_30.60cm_mean.1000 + clay_15.30cm_mean.2500 + latitude_decimal, data = df, REML = TRUE)
summary(hyp4)
```



### H5: lmer(LRR ~ (1|ma_id) + (1|study_id) + fieldsize, data = df) 
```{r}
hyp5 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + fieldsize, data = df) 
summary(hyp5)
```

#### predict 
```{r gow to use predict }
hyp5 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + fieldsize, data = df) 
lmer1 <- lmer(y ~ x + (1|unit), data = X)

X <- df %>% 
  mutate(fit.m = predict(hyp5, re.form = NA),
         fit.c = predict(hyp5, re.form = NULL))

X %>%
  ggplot(aes(x = fieldsize, y = LRR)) +
    geom_point(pch = 16, col = "black") +
    geom_line(aes(y = fit.m), col = 1, size = 2) +
    coord_cartesian(ylim = c(-40, 100))+
    theme_bw()


X <- X %>%
  mutate(resid = resid(hyp5))

X %>%
  ggplot(aes(x = fieldsize, y = fit.m + resid)) +
  geom_point(pch = 16) +
  geom_line(aes(y = fit.m), col = 1, size = 2)  +
  coord_cartesian(ylim = c(-10, 10))+
  facet_wrap(~crop_type_grouped_big) +
  theme_bw()
```



### H6: hyp6 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|fieldsize) + SR, data = df )

```{r}
dfhyp6 <- 
df %>% 
  select(LRR, ma_id, study_id, fieldsize, SR, crop_type_grouped_big) %>% 
  na.omit()

hyp6 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + SR, data = dfhyp6 )
summary(hyp6)
```

```{r}
X <- dfhyp6 %>% 
  mutate(fit.m = predict(hyp6, re.form = NA),
         fit.c = predict(hyp6, re.form = NULL))

X %>%
  ggplot(aes(x = SR, y = LRR)) +
    geom_point(pch = 16, col = "black") +
    geom_line(aes(y = fit.m), col = 1, size = 2) +
    coord_cartesian(ylim = c(-40, 100))+
    theme_bw()


X <- X %>%
  mutate(resid = resid(hyp6))

X %>%
  ggplot(aes(x = SR, y = fit.m + resid, col = crop_type_grouped_big)) +
  geom_point(pch = 16) +
  geom_line(aes(y = fit.m), col = 1, size = 1)  +
  coord_cartesian(ylim = c(-40, 100))+
  facet_wrap(~fieldsize) +
  theme_bw()
```


### H7: lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|control_id) + nat.hab.1000*fieldsize, data = df_dependent)  
```{r}
#hyp7 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + nat.hab.1000*fieldsize, data = df_dependent)  
```


### H8: hyp8 <- lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|control_id) + shannon.1000*fieldsize, data = df)

```{r}
# model 
dfhyp8 <- 
df %>% 
  select(LRR, ma_id, study_id, fieldsize, SR, shannon.1000, crop_type_grouped_big, treatment, poll_dependent) %>% 
  na.omit()

hyp8 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|poll_dependent) + shannon.1000*SR, data = dfhyp8 )
summary(hyp8)
AIC(hyp8)

#plot check
plot_model(hyp8, type = "slope") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 
```
```{r}

hyp8_eq <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|poll_dependent) + I(shannon.1000^2)*I(SR^2), data = dfhyp8 )
summary(hyp8_eq)
AIC(hyp8_eq)

#plot check
plot_model(hyp8_eq, type = "resid") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 

#plot check
plot_model(hyp8_eq, type = "pred", terms="shannon.1000 [all]") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 

```

```{r}
Cond_DF <- as.data.frame(ranef(hyp8_eq)) %>% transmute(unit = grp, b0_hat = condval) %>% mutate(Intercept_cond = b0_hat + summary(hyp8_eq)$coef[1,1])

X %>%
  ggplot(aes(x = shannon.1000, y = fit.m + resid)) +
  geom_point(pch = 16, col = "grey") +
  geom_violinhalf(data = Cond_DF, aes(x = 0, y = Intercept_cond), trim = FALSE, width = 3, fill = NA) +
  geom_line(aes(y = fit.m), col = 1, size = 2) +
  coord_cartesian(ylim = c(-1, 1))
```

```{r}
ggplot(df, aes(x = LRR, y = shannon.1000)) +
  geom_violin() +
  geom_point()+
  labs(x = "LRR", y = "shannon") +
  theme_minimal()
```

```{r}
# plot 
X <- dfhyp8 %>% 
  mutate(fit.m = predict(hyp8_eq, re.form = NA),
         fit.c = predict(hyp8_eq, re.form = NULL))

X %>%
  ggplot(aes(x = SR, y = LRR, col = I(shannon.1000^2))) +
    geom_point(pch = 16, col = I(shannon.1000^2)) +
    geom_line(aes(y = fit.m), col = 1, size = 2) +
    coord_cartesian(ylim = c(-40, 100))+
    theme_bw()


X <- X %>%
  mutate(resid = resid(hyp8))

X %>%
  ggplot(aes(x = SR, y = fit.m + resid, col = shannon.1000)) +
  geom_point(pch = 16) +
  geom_line(aes(y = fit.m), col = 1, size = 1)  +
  coord_cartesian(ylim = c(-25, 25))+
  facet_wrap(~poll_dependent) +
  theme_bw()
```

### H8.2
```{r}

hyp8_v2 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|poll_dependent) + ns(shannon.1000, df = 3)*ns(SR, df = 2), data = dfhyp8 )
summary(hyp8_v2)
AIC(hyp8_v2)

#plot check
plot_model(hyp8_v2, type = "resid") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 

#plot check
plot_model(hyp8_v2, type = "pred", terms="shannon.1000 [all]") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 

```

```{r}
hyp8_v2 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|poll_dependent) + ns(shannon.1000, df = 3)*ns(SR, df = 2), data = dfhyp8 )

X <- dfhyp8 %>% 
  mutate(fit.m = predict(hyp8_v2, re.form = NA),
         fit.c = predict(hyp8_v2, re.form = NULL))

X %>%
  ggplot(aes(x = shannon.1000, y = LRR, col = SR)) +
    geom_point(pch = 16) +
    geom_smooth(aes(y = fit.m), col = 1, size = 1) +
    coord_cartesian(ylim = c(-10, 10))+
    theme_bw()


X <- X %>%
  mutate(resid = resid(hyp8_v2))

X %>%
  ggplot(aes(x = shannon.1000, y = fit.m + resid)) +
  geom_point(pch = 16, size = 1.5) +
  geom_smooth(aes(y = fit.m), col = 1, size = 1)  +
  coord_cartesian(ylim = c(-25, 25))+
  facet_wrap(~poll_dependent) +
  theme_bw()
```



### H9

```{r}
# data 
dfhyp9 <- 
df %>% 
  select(LRR, ma_id, study_id, shannon.1000, crop_type_grouped_big, treatment, poll_dependent, cropland.1000,
        shannon.2500, shannon.5000,cropland.2500, cropland.5000) %>% 
  na.omit()


hyp9 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + ns(shannon.1000, df = 3)*ns(cropland.1000, df = 3), data = dfhyp9)
summary(hyp9)
AIC(hyp9)


#plot check
plot_model(hyp9, type = "slope", terms = c("shannon.1000", "cropland.1000")) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 
```

```{r}
# plot 
X <- dfhyp9 %>% 
  mutate(fit.m = predict(hyp9, re.form = NA),
         fit.c = predict(hyp9, re.form = NULL))

X %>%
  ggplot(aes(x = shannon.1000, y = LRR, col = cropland.1000)) +
    geom_point(pch = 16, col = cropland.1000) +
    geom_line(aes(y = fit.m), col = 1, size = 2) +
    coord_cartesian(ylim = c(-40, 100))+
    theme_bw()


X <- X %>%
  mutate(resid = resid(hyp9))

X %>%
  ggplot(aes(x = cropland.1000, y = fit.m + resid, col = shannon.1000)) +
  geom_point(pch = 16, size = 1.5) +
  geom_line(aes(y = fit.m), col = 1, size = 1)  +
  coord_cartesian(ylim = c(-10, 10))+
  facet_wrap(~crop_type_grouped_big) +
  theme_bw()
```

# NEW BEGINNING 2.0 

```{r}
hist(df$LRR)
hist(df$SR) # slightly right skewed
hist(df$cropland.1000)
hist(df$nat.hab.1000) # right skewed
hist(df$shannon.1000) 
hist(df$soc_15.30cm_mean.1000) # very much right skewed
hist(df$medium_area_m.1000) # very left skewed 
hist(df$prec_avg_1970.2000) 
hist(df$temp_avg_1970.2000) 
hist(df$LRR) 
hist(df$landcover_map_year) # slightly left skewed
plot(df$nat.hab.1000, df$LRR)
```

### H10: latitude_decimal

```{r}
dfhyp10 <- 
df %>% 
  select(LRR, ma_id, study_id, shannon.1000, crop_type_grouped_big, treatment, poll_dependent, cropland.1000,
        shannon.2500, shannon.5000,cropland.2500, cropland.5000, prec_avg_1970.2000, latitude_decimal, temp_avg_1970.2000) %>% 
  na.omit()


hyp10 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + ns(latitude_decimal, df = 4), data = dfhyp10)
summary(hyp10)
AIC(hyp10)


plot_model(hyp10, type = "slope", terms = "latitude_decimal") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 

X <- dfhyp10 %>% 
  mutate(fit.m = predict(hyp10, re.form = NA),
         fit.c = predict(hyp10, re.form = NULL))

X %>%
  ggplot(aes(x = latitude_decimal, y = LRR)) +
    geom_point(pch = 16) +
    geom_line(aes(y = fit.m), col = 1, size = 2) +
    coord_cartesian(ylim = c(-40, 100))+
    theme_bw()


X <- X %>%
  mutate(resid = resid(hyp10))

X %>%
  ggplot(aes(x = latitude_decimal, y = fit.m + resid)) +
  geom_point(pch = 16, size = 1.5) +
  geom_line(aes(y = fit.m), col = 1, size = 1)  +
  coord_cartesian(ylim = c(-25, 25))+
  #facet_wrap(~crop_type_grouped_big) +
  theme_bw()

```

### H 11
```{r}
dfhyp11 <- 
df %>% 
  select(LRR, ma_id, study_id, shannon.1000, crop_type_grouped_big, treatment, poll_dependent, cropland.1000,
        shannon.2500, shannon.5000,cropland.2500, cropland.5000, prec_avg_1970.2000, latitude_decimal, temp_avg_1970.2000, country_new) %>% 
  na.omit()

hyp11 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + prec_avg_1970.2000, data = dfhyp11)
summary(hyp11)
AIC(hyp11)


plot_model(hyp11, type = "slope", terms = "") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 

X <- dfhyp11 %>% 
  mutate(fit.m = predict(hyp11, re.form = NA),
         fit.c = predict(hyp11, re.form = NULL))

X %>%
  ggplot(aes(x = prec_avg_1970.2000, y = LRR)) +
    geom_point(pch = 16) +
    geom_line(aes(y = fit.m), col = 1, size = 2) +
    coord_cartesian(ylim = c(-40, 100))+
    theme_bw()


X <- X %>%
  mutate(resid = resid(hyp11))

X %>%
  ggplot(aes(x = prec_avg_1970.2000, y = fit.m + resid, col = ma_id)) +
  geom_point(pch = 16, size = 1.5) +
  geom_line(aes(y = fit.m), col = 1, size = 1)  +
  coord_cartesian(ylim = c(-25, 25))+
  facet_wrap(~country_new) +
  theme_bw()
```
```{r}
dfhyp12 <- 
df %>% 
  select(LRR, ma_id, study_id, shannon.1000, crop_type_grouped_big, treatment, poll_dependent, cropland.1000,
        shannon.2500, shannon.5000,cropland.2500, cropland.5000, prec_avg_1970.2000, latitude_decimal, temp_avg_1970.2000, country_new) %>% 
  na.omit()

hyp12 <-  lmer(LRR ~ (1|ma_id) + (1|study_id) + (1|country_new) + temp_avg_1970.2000, data = dfhyp11)
summary(hyp12)
AIC(hyp12)


plot_model(hyp12, type = "slope", terms = "") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title   = element_blank(),      
    plot.subtitle = element_text(size = 12)
  ) 

X <- dfhyp11 %>% 
  mutate(fit.m = predict(hyp12, re.form = NA),
         fit.c = predict(hyp12, re.form = NULL))

X %>%
  ggplot(aes(x = temp_avg_1970.2000, y = LRR)) +
    geom_point(pch = 16) +
    geom_line(aes(y = fit.m), col = 1, size = 2) +
    coord_cartesian(ylim = c(-40, 100))+
    theme_bw()


X <- X %>%
  mutate(resid = resid(hyp12))

X %>%
  ggplot(aes(x = temp_avg_1970.2000, y = fit.m + resid, col = ma_id)) +
  geom_point(pch = 16, size = 1.5) +
  geom_line(aes(y = fit.m), col = 1, size = 1)  +
  coord_cartesian(ylim = c(-25, 25))+
  facet_wrap(~treatment) +
  theme_bw()
```

### Manuscript 

1.	Land cover diversity (measured with the Shannon index) correlates positively with yield treatment effectiveness.
2.	The correlation between percentage of natural habitat and yield varies between crop types and treatments, due to pollinator dependence and edge effects.
3.	The correlation between percentage of natural habitat and yield varies with total species richness, climate and location, due to differences in biogeographical characteristics.
4.	The percentage of arable area is negatively associated with yield, but not in a linear way, because a steep increase of arable land in an area can lead to land degradation and decrease of species richness. 
5.	Landscape composition and configuration within 1 km radius from the field affect the effectiveness of yield treatments stronger than within 5 km radius.





### Clustering 
```{r}
# Select variables for clustering
cluster_vars <- df %>%
  select(LRR, SR, nat.hab.1000, shannon.1000,, cropland.1000, prec_avg_1970.2000, temp_avg_1970.2000, latitude_decimal) %>%
  mutate(across(where(is.matrix), as.vector)) %>%
  mutate(across(everything(), as.numeric)) %>%
  scale() # Important: standardize variables

# Remove NAs
cluster_vars <- na.omit(cluster_vars)

# Determine optimal number of clusters
library(factoextra)
fviz_nbclust(cluster_vars, kmeans, method = "wss") # Elbow method
fviz_nbclust(cluster_vars, kmeans, method = "silhouette") # Silhouette method

# Apply k-means with optimal k
set.seed(123)
k <- 3 # Choose based on elbow/silhouette plot
kmeans_result <- kmeans(cluster_vars, centers = k, nstart = 25)

# Add clusters back to original data (matching rows)
df_clean <- df[complete.cases(select(df, LRR, SR, nat.hab.1000, clay_0.5cm_mean.1000, soc_0.5cm_mean.1000)), ]
df_clean$cluster <- as.factor(kmeans_result$cluster)

# Visualize clusters
fviz_cluster(kmeans_result, data = cluster_vars, 
             palette = "Set2", ggtheme = theme_minimal(),
             main = "K-means Clustering of Study Sites")
```
### RDA 
```{r}
# package
library(vegan)
dff <- df %>% 
  select(LRR, nat.hab.1000, shannon.1000, cropland.1000, prec_avg_1970.2000, temp_avg_1970.2000, latitude_decimal)
dff <- df %>% 
  select(LRR, soc_15.30cm_mean.1000, nitrogen_15.30cm_mean.1000, bdod_15.30cm_mean.1000) %>% 
  drop_na()
res.pca <- prcomp(dff,  scale = TRUE)
fviz_pca_ind(res.pca, col.ind = dff$LRR)
```

you will have diffrent slopes with different field sizes -> NO! (like how and why i dont know, but there is no difference in slopes)
what do you expect?: 

grouping: 
- ma_id
- study_id 

varying grouping: 
- fieldsize -> dif tretments applied, dif edges
- treatments 
- crop_types
- pol dependent 

### H1 but nested random
```{r}
hyp1_nest <- lmer(LRR ~ (1|ma_id/study_id) + SR + nat.hab.peri.area.ratio.1000 + poll_dependent, data = df, REML = TRUE)

summary_hyp1_nest <- summary(hyp1_nest)
AIC_hyp1_nest <- AIC(hyp1_nest) # 3575.568

summary_hyp1 <- summary(hyp1)
AIC_hyp1 <- AIC(hyp1) # 3644.524

```



### Findings 

- increase in dispersion of natural habitat is increasing the success of the treatment (nat.hab.peri.area.ratio.1000) (hyp1)












